### 2021-06-17 paper/model/summary.tex
Got started re-writing with Radost's comments on the side.

The structure of the model is depicted in Figure~\ref{fig:broad_model_description}.

The background characteristics include age, county and occupation of each simulated
individual. Contact models are functions that map individual characteristics into a
predicted number of contacts. Currently, we distinguish between eight types of contact
models which are all listed in Figure~\ref{fig:broad_model_description}: households,
recurrent and random work contacts, recurrent and random leisure contacts, and nursery,
preschool, and school contacts.

@@

Once a person is infected, the disease progresses in a fairly standard way which is
depicted in Figure~\ref{fig:disease_progression}. Asymptomatic cases and cases with mild
symptoms are infectious for some time and recover eventually. Cases with severe symptoms
additionally require hospitalization and lead to either recovery or death.

>>>>

The structure of the model is depicted in Figure~\ref{fig:model_contacts_infections}.
\comment[id=HM]{Mentioned by Tim in the course, just main paper: Modular structure may deserve a sentence}

The background characteristics include age, county and occupation of each simulated
individual. Contact models are functions that map individual characteristics into a
predicted number of contacts. Currently, we distinguish between eight types of contact
models which are all listed in Figure~\ref{fig:model_contacts_infections}: households,
recurrent and random work contacts, recurrent and random leisure contacts, and nursery,
preschool, and school contacts.

@@

Once a person is infected, the disease progresses in a fairly standard way which is
depicted in Figure~\ref{fig:model_disease_progression}. Asymptomatic cases and cases with mild
symptoms are infectious for some time and recover eventually. Cases with severe symptoms
additionally require hospitalization and lead to either recovery or death.

### 2021-06-17 paper/report.tex
Got started re-writing with Radost's comments on the side.

Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
presented an enormous challenge to humanity on many dimensions. The development of highly
effective vaccines holds the promise of containment in the medium term. However, most
countries find themselves many months---and often years---away from reaching
vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
some more scientific citation before submission} In the meantime, it is of utmost
importance to employ an effective mix of strategies for containing the virus. The most
frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
contacts between individuals. While this has allowed some countries to sustain equilibria
with very low infection numbers\footnote{See \citet{Contreras2021} for a theoretical
equilibrium at low case numbers which is sustained with test-trace-and-isolate
policies.}, most have seen large fluctuations of infection rates over time. Containment
measures have become increasingly diverse and now include testing, more nuanced NPIs, and
contact tracing. Neither these policies' effect nor the influence of seasonal patterns or
more infectious virus strains are well understood in quantitative terms.

This paper develops a quantitative model incorporating all these factors. The framework
allows to combine a wide variety of data and mechanisms in a timely fashion, making it
useful to predict the effects of various interventions. We apply the model to Germany
and show that rapid testing had the largest impact on the reduction in infections by
almost 80\% during the month of May 2021. We conclude that rapid tests have a large role
to play for at least as long as vaccinations have not been offered to an entire
population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:broad_model_description}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
model most closely related to ours is described in \citet{Hinch2020}.} Each contact
between an infectious individual and somebody susceptible to the disease bears the risk
of transmitting the virus. Contacts occur in the household, at work, at school, or in
other settings (leisure activities, grocery shopping, medical appointments, etc.). Some
contacts recur regularly, others occur at random. Random contacts are typically
assortative in age and geographical location. Empirical applications can take the
population structure from census data and the types and frequency of contacts from diary
data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008,Hoang2019}.

The three dimensions of contact types are chosen so that the most common NPIs can be
modeled in great detail. NPIs affect the number of contacts or the risk of transmitting
the disease upon having physical contact. The effect of different NPIs will generally
vary across contact types. For example, a mandate to work from home will reduce the
number of work contacts to zero for a fraction of the working population. Schools and
daycare can be closed entirely, operate at reduced capacity---including an alternating
schedule---,or implement mitigation measures like masking requirements or air filters
\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
settings. In any setting, measures like masking requirements would reduce the
probability of infection associated with a contact \citep{Cheng2021}.

@@

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
        \caption{{Model description}}
        \label{fig:broad_model_description}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
        \vskip5ex

        \caption{Disease progression}
        \label{fig:disease_progression}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering

        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
        \caption{{Model for antigen tests}}
        \label{fig:antigen_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
        \vskip5ex

        \caption{{Translating all infections to recorded ones}}
        \label{fig:model_for_official_cases}
    \end{subfigure}

    \caption{Model description}
    \label{fig:model-description}

    \floatfoot{\noindent \textit{Note:}
        A
        % formal 
        description of the model can be found in Supplementary Material~\ref{sec:model}.
        Figure~\ref{fig:broad_model_description} shows the influence on an agent's
        contacts to other agents. Demographic characteristics set the baseline number of
        contacts in different networks. She may reduce the number of contacts due to
        NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
        occur when a susceptible agent meets an infectious agent; the probability
        depends on the type of contact, on seasonality, and on NPIs. If infected, the
        infection progresses as depicted in Figure~\ref{fig:disease_progression}. If
        rapid tests are avaible, agents' demand is modeled as in
        Figure~\ref{fig:antigen_tests}. All pupils are attending school are tested,
        other reasons trigger a test only for a fraction of individuals.
        Figure~\ref{fig:model_for_official_cases} shows the model of translating all
        infections in the simulated data to age-specific recorded infections. The model
        uses data on the aggregate share of recorded cases, on the share of positive PCR
        tests triggered by symptoms, and on the false positive rate of rapid tests. The
        lower part of the graph is relevant only for periods where rapid tests are
        available.}
\end{figure}

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
The probabilities of transitioning between these states depend on age; their duration is
random within intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
some days to be processed and there are aggregate capacity constraints throughout. In
contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
that are demanded will be performed. Specificity and sensitivity of these tests is set
according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
timing of the test relative to the start of infectiousness.
Figure~\ref{fig:antigen_tests} shows our model for rapid test demand. Schools may require
students to be tested regularly. Rapid tests may be offered by employers for on-site
workers. Individuals may demand tests for private reasons, which include having plans to
meet other people\footnote{A positive test will make them reduce their contacts; this is
why tests impact the actual contacts in Figure~\ref{fig:model-description}.}, showing
symptoms of CoViD-19, and because a household member tested positively for the virus. We
endow each agent with an individual compliance parameter. This parameter determines
whether she takes up rapid tests offered by employers or follows up on private reasons
and whether she reduces her contacts if she tests positive or develops symptoms. The
thresholds fall over time and are calibrated to survey data where possible (see
Supplementary Materials~\ref{subsec:rapid_test_demand}).
% The thresholds are higher for tests in a private setting than for tests at the
% workplace until mid April and then compliance because of private reasons becomes higher
% than compliance at the workplace.

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group. Mobility data is used
to model the reductions in work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests.

@@

We estimate these parameters such that the model matches the following empirical
quantities: 1. official case numbers for each age group and region, 2. deaths, 3. share
of B.1.1.7. \citep[See][for the general method]{McFadden1989}. However, only a fraction
of infections actually enter the official case numbers. We thus model official cases as
depicted in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of the
share of detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As
the share of asymptomatic individuals varies by age group, this gives us age-specific
shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases
by age group over time in our model). Our estimates suggest that only 18 to 28\% of cases
are detected for school age children while the rate is 30 to 55\% for those above age
80.\comment[id=K]{Must be verified once results are final.} Confirmation of a positive
rapid test result is another reason for PCR tests, once these become available.

The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. We transform the index so that lower values represent
higher levels of restrictions. A value of zero means all measures incorporated in the
index are turned on. The value 1 represents the situation in mid-September, with
restrictions on gatherings and public events, masking requirements, but open schools and
workplaces. In the seven weeks between mid September and early November, cases increased
by a factor of 10. Restrictions were somewhat tightened in mid-October and again in early
November. New infections remained constant throughout November, before rising again in
December, which prompted the most stringent lockdown to this date. Schools and daycare
centers were closed again, so were customer-facing businesses except for grocery and drug
stores. From the peak of the second wave just before Christmas until the trough in
mid-February, newly detected cases decreased by almost three quarters. The third wave in
the spring of 2021 is associated with the B.1.1.7 strain, which became dominant in March.
See Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
stringency index.

@@

The largest effect---almost one half when considering the decompositions---comes from
rapid testing. Here, it is crucial to differentiate between recorded cases and actual
cases. Additional testing means that infections are detected, which would otherwise
remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
until late April, recorded cases are higher than in the scenario where none of the three
mechanisms are turned on. Compared to the scenario with vaccinations only, this point is
reached around mid-May and it would be June for the comparison with the seasonality-only
scenario. The effect on total cases, however, is visible immediately. Despite the fact
that only a small fraction of the population performed weekly rapid tests in
March\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
week of March. 8\% in the model have ever done a rapid test at that point; 27\% of
respondents to the COSMO study reported in early March that they \textit{ever} did a
rapid test)}, new infections on April 1 would be reduced by 53\% relative to the
scenario without vaccinations, rapid tests, or seasonality.

>>>>

Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
presented an enormous challenge to humanity on many dimensions. The development of highly
effective vaccines holds the promise of containment in the medium term. However, most
countries find themselves many months---and often years---away from reaching
vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
some more scientific citation before submission} In the meantime, it is of utmost
importance to employ an effective mix of strategies for containing the virus. The most
frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
contacts between individuals. While this has allowed some countries to sustain equilibria
with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
equilibrium at low case numbers which is sustained with test-trace-and-isolate
policies.} most have seen large fluctuations of infection rates over time. Containment
measures have become increasingly diverse and now include testing, more nuanced NPIs, and
contact tracing. Neither these policies' effect nor the influence of seasonal patterns or
more infectious virus strains are well understood in quantitative terms.

This paper develops a quantitative model incorporating these factors simultaneously. The
framework allows to combine a wide variety of data and mechanisms in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany and show that rapid testing had the largest impact on the reduction in
infections by almost 80\% during the month of May 2021. We conclude that rapid tests
have a large role to play for at least as long as vaccinations have not been offered to
an entire population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
model most closely related to ours is described in \citet{Hinch2020}.} Each contact
between an infectious individual and somebody susceptible to the disease bears the risk
of transmitting the virus. Contacts occur in up to four networks: Within the household,
at work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, contacts are
differentiated by age and geographical location see Section~\ref{sub:number_of_contacts}
of the Supplementary Material.

The four contact networks are chosen so that the most common NPIs can be
modeled in great detail. NPIs affect the number of contacts or the risk of transmitting
the disease upon having physical contact. The effect of different NPIs will generally
vary across contact types. For example, a mandate to work from home will reduce the
number of work contacts to zero for a fraction of the working population. Schools and
daycare can be closed entirely, operate at reduced capacity---including an alternating
schedule---, or implement mitigation measures like masking requirements or air filters
\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
settings. In any setting, measures like masking requirements would reduce the
probability of infection associated with a contact \citep{Cheng2021}.

@@

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
        \caption{{Model for contacts and infections}}
        \label{fig:model_contacts_infections}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
        \vskip5ex

        \caption{Disease progression}
        \label{fig:model_disease_progression}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering

        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
        \caption{{Model for rapid tests}}
        \label{fig:model_rapid_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
        \vskip5ex

        \caption{{Translating all infections to recorded ones}}
        \label{fig:model_official_cases}
    \end{subfigure}

    \caption{Model description}
    \label{fig:model_description}

    \floatfoot{\noindent \textit{Note:}
        A
        % formal 
        description of the model can be found in Supplementary Material~\ref{sec:model}.
        Figure~\ref{fig:model_contacts_infections} shows the influence on an agent's
        contacts to other agents. Demographic characteristics set the baseline number of
        contacts in different networks. She may reduce the number of contacts due to
        NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
        occur when a susceptible agent meets an infectious agent; the probability
        depends on the type of contact, on seasonality, and on NPIs. If infected, the
        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}. If
        rapid tests are avaible, agents' demand is modeled as in
        Figure~\ref{fig:model_rapid_tests}. All pupils are attending school are tested,
        other reasons trigger a test only for a fraction of individuals.
        Figure~\ref{fig:model_official_cases} shows the model of translating all
        infections in the simulated data to age-specific recorded infections. The model
        uses data on the aggregate share of recorded cases, on the share of positive PCR
        tests triggered by symptoms, and on the false positive rate of rapid tests. The
        lower part of the graph is relevant only for periods where rapid tests are
        available.}
\end{figure}

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
The probabilities of transitioning between these states depend on age; their duration is
random within intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
some days to be processed and there are aggregate capacity constraints throughout. In
contrast, rapid antigen tests yield immediate results; after a phase-in period, all
tests that are demanded will be performed. Specificity and sensitivity of these tests is
set according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on
the timing of the test relative to the onset of infectiousness.
Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
require students to be tested regularly. Rapid tests may be offered by employers for
on-site workers. Individuals may demand tests for private reasons, which include having
plans to meet other people,\footnote{A positive test will make them reduce their
contacts; this is why tests impact the actual contacts in
Figure~\ref{fig:model_description}.} showing symptoms of CoViD-19, and because a
household member tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid tests offered
by employers or follows up on private reasons.\comment[id=HM]{Removed ``and whether she
reduces her contacts if she tests positive or develops symptoms'' -- does not fit here.
If anything, add footnote, but it seems like this is a detail that is just distracting
right here. Make sure to describe clearly and at length in Appendix.}

% The thresholds fall over time and are calibrated to survey data where possible (see
% Supplementary Materials~\ref{subsec:rapid_test_demand}).

% The thresholds are higher for tests in a private setting than for tests at the
% workplace until mid April and then compliance because of private reasons becomes
% higher than compliance at the workplace.

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group. Mobility data is used
to model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests.

@@

We estimate these parameters such that the model matches the following empirical
quantities: 1. official case numbers for each age group and region, 2. deaths, 3. share
of B.1.1.7. \citep[See][for the general method]{McFadden1989}. However, only a fraction
of infections actually enter the official case numbers. We thus model official cases as
depicted in Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the
share of detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As
the share of asymptomatic individuals varies by age group, this gives us age-specific
shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases
by age group over time in our model). Our estimates suggest that only 18 to 28\% of cases
are detected for school age children while the rate is 30 to 55\% for those above age
80.\comment[id=K]{Must be verified once results are final.} Confirmation of a positive
rapid test result is another reason for PCR tests, once these become available.

The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. For legibility, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value 1 represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of 10. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November, before rising again in December, which prompted the most stringent lockdown to
this date. Schools and daycare centers were closed again, so were customer-facing
businesses except for grocery and drug stores. From the peak of the second wave just
before Christmas until the trough in mid-February, newly detected cases decreased by
almost three quarters. The third wave in the spring of 2021 is associated with the
B.1.1.7 strain, which became dominant in March. See Figure~\ref{fig:share_b117}. In
early March, some NPIs were being relaxed; e.g., hairdressers and home improvement
stores were allowed to open again to the public. There were many changes in details of
regulations afterwards, but they did not change the stringency index.

@@

The largest effect---almost one half when considering the decompositions---comes from
rapid testing. Here, it is crucial to differentiate between recorded cases and actual
cases. Additional testing means that infections are detected, which would otherwise
remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
until late April, recorded cases are higher than in the scenario where none of the three
mechanisms are turned on. Compared to the scenario with vaccinations only, this point is
reached around mid-May and it would be June for the comparison with the seasonality-only
scenario. The effect on total cases, however, is visible immediately. Despite the fact
that only a small fraction of the population performed weekly rapid tests in
March,\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
week of March. 8\% in the model have ever done a rapid test at that point; 27\% of
respondents to the COSMO study reported in early March that they \textit{ever} did a
rapid test)} new infections on April 1 would be reduced by 53\% relative to the
scenario without vaccinations, rapid tests, or seasonality.
